---
title: "One_Variable_R"
output: html_document
---
### Reading in Data
```{r}
getwd()
list.files()
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
names(pf)
```   
#Plot histogram of day of Birth
```{r}
if (!require('ggplot2'))
{
install.packages("ggplot2")
library(ggplot2)
}

qplot(x= dob_day, data= pf)+
  scale_x_continuous(breaks=1:31)+
  facet_wrap(~dob_month)
```

#plot histogram of frient count
# Will load pf again as its required
```{r}
library(ggplot2)
qplot(x=friend_count, data=pf, binwidth=25)+
  scale_x_continuous(limits= c(0,1000),breaks = seq(0,1000,50))+
  facet_wrap(~gender)

``` 
#Trying the same using ggplot
```{r}
ggplot(aes(x = friend_count), data = pf) +
  geom_histogram()

```

#Omitting NA from the Plot
```{r}
qplot(x=friend_count, data = subset(pf, !is.na(gender)), binwidth= 25)+
  scale_x_continuous(limits=c(0,1000), breaks = seq(0,1000,50))+
  facet_wrap(~gender)

``` 
#Statistics using table, and by 
```{r}
table(pf$gender)
by(data=pf$friend_count, pf$gender, summary)
```
#Histogram of tenure in years with appropriate bin width

  

```{r}
qplot(x=(tenure)/365, data=pf, binwidth=0.25,   color = I('black'), fill = I('#F79420'), xlab = "No of years in FB", ylab = 'No of users')+
  scale_x_continuous(breaks=seq(1,7,1), limits = c(0,7))
```


#Histogram of Ages
```{r}
qplot(x=age, data=pf, binwidth=1,   color = I('black'), fill = I('#F79420'), xlab = "Age of Users", ylab = 'No of users')+
  scale_x_continuous(limits = c(10,70), breaks = seq(0,70,5))
```
#Friend count 
#Transformed using log 10
#Transformed using sqrt
```{r}
if (!require(gridExtra))
{
install.packages('gridExtra')
}
library(gridExtra)
  
library(ggplot2)

p1 = qplot(x=friend_count, data = pf,color = I('black'), fill = I('#F79420'), xlab = 'Friend Count(normal Scale)', ylab = 'No of Users')

p2 = qplot(x=log(friend_count+1), data = pf,color = I('black'), fill = I('#F79420'), xlab = 'Friend Count(log Scale)', ylab = 'No of Users')


p3 = qplot(x=sqrt(friend_count), data = pf,color = I('black'), fill = I('#F79420'), xlab = 'Friend Count(squre Scale)', ylab = 'No of Users')
  
grid.arrange(p1, p2, p3, ncol=2)

```
#Using the same using ggplot 

```{r}

p4 = ggplot(aes(x = friend_count), data= pf) + geom_histogram()
p5 = p1+scale_x_log10()
p6 = p1+scale_x_sqrt()

grid.arrange(p4,p5,p6, ncol=3)

```
#Frequency Polygon of gets more likes on world wide web 

```{r}

qplot(x = www_likes, data = subset(pf, !is.na(gender)),geom = 'freqpoly', color = gender)+
  scale_x_continuous()+
  scale_x_log10()




```

#Who has more likes 

```{r}

by(pf$www_likes,pf$gender, sum)
```
#Box Plots
```{r}
qplot(x=gender, y=friend_count, geom = 'boxplot', data= subset(pf, !is.na(gender)), ylim = c(0,1000))

#But this removed the values, so a better way to do this to use coord cartesian

qplot(x=gender, y=friend_count, geom='boxplot', data= subset(pf,!is.na(gender))) + coord_cartesian(ylim=c(0,1000))

# Zooming further 
qplot (x=gender, y=friend_count, geom='boxplot', data=subset(pf ,!is.na(gender))) + coord_cartesian(ylim= c(0,250))

by(pf$friend_count, pf$gender, summary)

```

#Logical variable creation 

```{r}

summary (pf$mobile_likes)
summary(pf$mobile_likes >0)

pf$mobile_checkin <- NA
pf$mobile_checkin <- ifelse(pf$mobile_likes>0, 1, 0)
pf$mobile_checkin <- factor(pf$mobile_checkin)

summary(pf$mobile_checkin)
#percentage
sum(pf$mobile_checkin == 1)/length(pf$mobile_checkin)

```
#For future