---
title: "EDA_DiamondDataSet"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load Library
```{r}
library(ggplot2)
ds = data(diamonds)
diamonds
str(diamonds)
str(unique(diamonds$color))
?(diamonds)

```

#Price Histogram
```{r}
qplot(x=price, data=diamonds, xlab = "Price", ylab = "No of Diamonds",color = I('black'), fill = I('#F79420')) 
  
summary(diamonds$price)
```
#Diamonds costing various prices
```{r}
sum (diamonds$price<500)
sum (diamonds$price<250)
sum (diamonds$price>=15000)
```

#Exploring
```{r}
qplot(x=price, data=diamonds, xlab = "Price", ylab = "No of Diamonds",color = I('black'), fill = I('#F79420'), binwidth = 500)+
  scale_x_continuous(limits = c(0,10000), breaks = seq(0,10000,500))
```
## Break out the histogram of diamond prices by cut.

```{r}
if (!require(gridExtra))
{
install.packages('gridExtra')
}
library(gridExtra)
  
library(ggplot2)

qplot(x=price, data=diamonds, xlab = "Price", ylab = "No of Diamonds",color = I('black'), fill = I('#F79420'))+
  facet_wrap(~cut)

by(diamonds$cut, data = diamonds$price, summary)
by(diamonds$cut, data = diamonds$price, max)

```
#Free up scales based on the numbers 

```{r}
hp <- ggplot(diamonds, aes(x=price)) + geom_histogram(colour="white", binwidth = 500)
hp + facet_wrap(~cut, scales='free')
```

# Create a histogram of price per carat
# and facet it by cut. You can make adjustments
# to the code from the previous exercise to get
# started.

# Adjust the bin width and transform the scale
# of the x-axis using log10.

```{r}
price_per_carat = ggplot(diamonds, aes(x=(price+1/carat))) +geom_histogram(color = I('black'), fill = I('#F79420'), binwidth = 1000)+xlab("Price Per Carat") 
price_per_carat + facet_wrap(~cut)

```

# Investigate the price of diamonds using box plots,
# numerical summaries, and one of the following categorical
# variables: cut, clarity, or color.

```{r}
qplot(x=cut, y=price, geom = 'boxplot', data= diamonds) 
qplot(x=clarity, y=price, geom = 'boxplot', data= diamonds)+ coord_cartesian(ylim= c(0,6000))
qplot(x=color, y=price, geom = 'boxplot', data= diamonds)

by( diamonds$price, diamonds$color, summary)

IQR(subset(diamonds, diamonds$color == 'D')$price)
IQR(subset(diamonds, diamonds$color == 'J')$price)

```

# Investigate the price per carat of diamonds across
# the different colors of diamonds using boxplots.
```{r}

qplot(x=color, y=(price/carat), geom = 'boxplot', data= diamonds)+ coord_cartesian(ylim=c(0,8000))

```
#Investigating weights of diamonds(carat) using a frequency Polygon
#Used Different Bin Widths to show how a frequency polygon changes

```{r}
library(gridExtra)
str(diamonds$carat)
p1 = qplot(x = carat, data = diamonds ,geom = 'freqpoly')+scale_x_continuous(breaks = seq(0,6,0.5))
p2= qplot(x = carat, data = diamonds ,geom = 'freqpoly', binwidth=0.5)+scale_x_continuous(breaks = seq(0,6,0.5))
p3= qplot(x = carat, data = diamonds ,geom = 'freqpoly', binwidth=0.1)+scale_x_continuous(breaks = seq(0,6,0.5))
grid.arrange(p1,p2,p3)


```
#Using GapMinder Data
#You might use a simple histogram, a boxplot split over a categorical variable,
#or a frequency polygon. The choice is yours!
```{r}

if(!require(readxl)){
  install.packages("readxl")
}
if(!require(tidyr)){
  install.packages("tidyr")
}
if(!require(dplyr)){
  install.packages("dplyr")
}

p_data <- read_excel('indicator_median age.xlsx')

#gathering years in Columns
gathered_p_data = tidyr::gather(p_data, "year","Median_Age", 2:22)

#remove any NA
cleaned_p_data = dplyr::filter(gathered_p_data, !is.na(Median_Age))
colnames(cleaned_p_data)[1] <- "Country"

#Choose Countries 
selected_countries <- c('India', 'United States', 'United Kingdom', 'China', 'Australia')

library(ggplot2)
ggplot(subset(cleaned_p_data, as.numeric(year)%%10 ==0 & Country %in% selected_countries))+
  geom_line(aes(x=year, y=Median_Age, color=Country, group = Country))+
  theme(legend.position="top") +
  labs(x='Year', y='Median Age')
  
```
#Box Plot of Population to give an example of absolute values 

```{r}
ggplot(subset(cleaned_p_data, Country %in% selected_countries), 
       aes(x=Country, y=Median_Age, fill = Country)) +
  geom_boxplot() +
  theme(legend.position="none", axis.text.x=element_text(angle=90)) +
  labs(x='Country', y='Median Age')
```

#Oil Production Example
```{r}

library(readxl)
xlsx_oil <- read_excel('Oil Production.xls.xlsx')
View(xlsx_oil)


library(tidyr)
library(dplyr)

tidy_oil <- tidyr::gather(xlsx_oil, "year", "production", 2:48)

oil_production <- dplyr::filter(tidy_oil, !is.na(production))

colnames(oil_production)[1] <-'Country'
#Try to plot only for a subset of countries 

selected_countries <- c('Brazil', 'Canada', 'China', 'Iraq', 'Nigeria', 'Russia', 'Saudi Arabia', 'United States', 'Venezuela')

library(ggplot2)
ggplot(subset(oil_production, as.numeric(year)%%5 == 0 & Country %in% selected_countries)) +
  geom_line(aes(x=year, y=production, colour=Country,  group=Country)) +
  theme(legend.position="top") +
  labs(x='Year', y='Oil Production')

```

#Shows the gaps between countries in terms of Oil Production 
```{r}
ggplot(subset(oil_production, Country %in% selected_countries), 
       aes(x=Country, y=production, fill = Country)) +
  geom_boxplot() +
  theme(legend.position="none", axis.text.x=element_text(angle=90)) +
  labs(x='Country', y='Oil Production')
```

# Your task is to investigate the distribution of your friends'
# birth months and days.
# Here some questions you could answer, and we hope you think of others.
# **********************************************************************

# How many people share your birthday? Do you know them?
# (Reserve time with them or save money to buy them a gift!)
# Which month contains the most number of birthdays?
# How many birthdays are in each month?
# Which day of the year has the most number of birthdays?
# Do you have at least 365 friends that have birthdays on everyday
# of the year?

```{r}

birthday_data <- read.csv('birthdaysExample.csv', header = T , stringsAsFactors = F)
birthday_data$dates.d <- as.Date(birthday_data$dates, format = "%m/%d/%y")


if (!require(lubridate)){
  install.packages('lubridate')
}

library(lubridate)


# How many people share your birthday? Do you know them?
length(birthday_data[month(birthday_data$dates.d)==09 & day(birthday_data$dates.d) ==27,])

# Which month contains the most number of birthdays?
library(ggplot2)

qplot(x = month(birthday_data$dates.d), data = birthday_data)+ scale_x_discrete( breaks = seq(1,12,1) )
#March

#How many birthdays are in each month?

count <- as.data.frame(table(month(birthday_data$dates.d)))
count[order(-count$Freq),]

#Which day of the year has the most birthdays?
count <- as.data.frame(table(yday(birthday_data$dates.d)))
count[order(-count$Freq),]
#37,142,197

#Do you have at least 365 friends that have birthdays on everyday
# of the year?

julian_days <- seq(1,365, 1)
missing <- julian_days[!(julian_days %in% count$Var1)]
ref = as.Date("2014-12-31")
for (i in 1:length(missing)){
  print(format(ref + missing[i], "%m/%d"))
}

```

